DOM bridges and barriers: systematic review and survey

### 2025-09-20 reports/draft_results.qmd
res fig 2

---
title: "DOM bridges and barriers: systematic review and end-user survey"
author: "Emily Nazario | UC Santa Cruz | enazario'@'ucsc.edu"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
editor_options: 
  chunk_output_type: console
---

@@

# Literature Review

## Selecting the literature that was included in this systematic review

In September 2024, a systematic search was completed using Web of Science to identify articles that discussed dynamic ocean management (DOM). Articles or documents published after this date were not included in this review. Articles were selected if they described factors contributing to the “success” of DOM, also termed here as a “bridge”, or that noted factors that should be addressed prior to using DOM approaches or support tools, also termed here as a “barrier”. In this systematic review, we coded bridges and barriers that were related to implemented or recommended DOM voluntary programs and/or enforced regulations, as well as those that were related to tool development (e.g., species spatial modeling tools such as EcoCast) that support dynamic management approaches.

During the initial search, the following table of terms were used. The "NEAR/5" operator was used to refine the search and ensure each of the terms in a given character string were found within one word of each other:

| Search term                |
|----------------------------|
| Dynamic ocean management   |
| Dynamic spatial management |
| Adaptive ocean management  |
| Dynamic ocean closures     |

: **Table 1.** List of terms used in the Web of Science search. Where spaces are found between words in a search term, the "NEAR/5" operator was added.

Works from the primary (i.e., peer-reviewed journals) and grey literature (e.g., theses, book chapters) were included, and this initial search resulted in 292 articles. Abstracts were skimmed for relevance, and a total of 141 articles were saved for further inspection An additional search using the NOAA Federal Register was also completed. Here, documents such as policy reports and planning documents were selected. This search resulted in 483 documents. The abstracts or content summaries for each of these documents were skimmed for relevance, and an additional 10 documents were saved for a full read through. Articles were excluded if DOM was mentioned or recommended, but there was no discussion as to why. Articles were also excluded if they discussed changes or updates to environmental data sources or remote sensing tools used to develop models supporting DOM, but DOM approaches were not specifically applied, recommended, or reviewed. Specific inclusion and exclusion criteria are described here:

| Inclusion | Exclusion |
|------------------------------------|:-----------------------------------|
| All cases where adaptive or near-real time forms of management or tools that support this form of management were discussed, even if the specific term 'Dynamic Ocean Management' was not explicitly used. | The term 'DOM' was mentioned, or even recommended, but there was no discussion as to why (e.g. SDM studies that suggest DOM applications as a next step for their model) |
| All cases where near real-time SDM tools or fleet-wide communication to avoid negative species interactions. | Articles that involved changes or updates to environmental data stores or remote sensing tools used for developing SDM tools to support DOM, but DOM was not specifically applied, recommended, or reviewed. |

The 151 articles were read by two reviewers and went through a final pass for determining inclusion. Articles that were referenced that seemed to discuss relevant topics were skimmed and their inclusion was determined. The final list of works analyzed for the review included 118 articles (i.e., primary literature, student theses, policy reports, and textbook chapters).

## Coding the bridges and barriers as identified in the literature

Two reviewers went through each of the 118 articles to identify text that was related to DOM bridges that were linked to the approach's success or were used as justification as to why DOM would be a good fit for that system/conservation problem. The reviewers also identified the barriers that hindered the efficacy of DOM or must be considered prior to its application. Through this effort, 19 “bridge” and 10 “barrier” categories were determined in the literature. “Bridge” and “barrier” factors were organized into four broader theme groups: ecological, social, and data. A table of the full list of codes/themes from the literature are depicted here. Please note that this is a draft of codes and that the sunburst diagrams need to be revised. I am also happy to modify the code/factor terminology throughout the study review process.

::: panel-tabset
### Bridges

### Barriers

![](images/barriers_old.png){fig-align="center" width="70%"}
:::

### Sankey diagram of literature code heirarchies

Here, each of the codes and their subcodes are presented. The presented percentages represent totals across all codes and codes closer to the top are codes that make up higher percentages of identified codes.

@@

  xlim(0.89, 3.6) +
  ylim(-1, 100)+
  labs(y = "Percent of all codes (%)", x = "") +
  theme_minimal()+
  theme(axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16), 
        legend.position = "none", 
        axis.text.x = element_blank(), 
        panel.grid.major = element_blank())+
  scale_fill_manual(values = c("#F48136", "#F46036", "#F44336", "#CFE2F3", "#C5E0B4", "#45818E"))+
  scale_color_manual(values = c("#F48136", "#F46036", "#F44336", "#CFE2F3", "#C5E0B4", "#45818E"))

ggsave(here("figures/sankey_tp.svg"), height = 10, width = 10, units = c("in"))

@@

## Visualizing the literature bridges and barrier codes

Below is a first attempt at how to visualize some of the data collected from this review. This first section represents general information about the articles selected for this review (e.g., study location, fisheries or shipping channels involved, species for which the tool or management program was established for). Most of these plots may instead go in the supplemental materials or aid with interpretation.

### Study information

#### Article categories
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 4

codes_vars$lit_type <- factor(codes_vars$lit_type , levels = c("NOAA Rule", "NMFS Notice", "Book chapter", "Thesis", "Methods", "Perspective", "Review", "Study"))

ggplot(codes_vars, aes(x = "", fill = lit_type))+
  geom_bar(position = "fill")+
  xlab("")+
  theme_minimal()+
  labs(fill = "Literature category")+
  ylab("Percentage of articles")+
  coord_flip()+
  scale_y_continuous(labels = scales::percent)+
  scale_fill_manual(values = MetBrewer::met.brewer("Veronese", n = 8))+
  theme(panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_blank(),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent'))
#ggsave(here("figures/lit_cat.svg"), width = 8, height = 4, units = c("in"))

::: panel-tabset
#### Study regions

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 7

codes_vars$broad_location <- as.factor(codes_vars$broad_location)
codes_LU <- codes_vars[!duplicated(codes_vars$Lit_unit), ]
reg_s <- codes_LU %>%
  group_by(broad_location) %>%
  filter(broad_location != "NA" & broad_location != "Europe and United States") %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(perc = count/sum(count)*100)

reg_bar <- ggplot(reg_s) +
  geom_bar(aes(x = perc, y = reorder(broad_location, perc)), fill = "dodgerblue4", width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     expand = c(0,0), 
                     limits = c(0, 48))+
  scale_y_discrete(expand = expansion(add = c(0.5,0.5))) + 
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 12)) + 
  geom_text(aes(x = perc, y = broad_location, label = broad_location),
    hjust = 0,
    nudge_x = 0.3,
    colour = "dodgerblue4",
    family = "Econ Sans Cnd",
    size = 3.5)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 13)
  )

reg_bar
```

#### Study fisheries and shipping channels

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-height: 10
#| fig-width: 10
fsh_shp <- codes_LU %>%
  gather(col, fsh_shp, fishery_or_channel1:fishery_or_channel4) %>%
  subset(select = fsh_shp)

fsh_shp_s <- fsh_shp %>%
  group_by(fsh_shp) %>%
  filter(fsh_shp != "NA") %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(perc = count/sum(count)*100)

fsh_shp_bar <- ggplot(fsh_shp_s) +
  geom_bar(aes(x = perc, y = reorder(fsh_shp, perc)), fill = "dodgerblue4", width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     expand = c(0,0), 
                     limits = c(0, 14))+
  scale_y_discrete(expand = expansion(add = c(0.5,0.5))) + 
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 15)) + 
  geom_text(aes(x = perc, y = fsh_shp, label = fsh_shp),
    hjust = 0,
    nudge_x = 0.3,
    colour = "dodgerblue4",
    family = "Econ Sans Cnd",
    size = 5)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 16)
  )

fsh_shp_bar
```

#### Study taxa

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 7

species <- codes_LU %>%
  gather(col, taxa, focal_taxa1:focal_taxa3) %>%
  subset(select = taxa)

spp_s <- species %>%
  group_by(taxa) %>%
  filter(taxa != "NA" & taxa != "Marine mammals") %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(perc = count/sum(count)*100)

taxa_bar <- ggplot(spp_s) +
  geom_bar(aes(x = perc, y = reorder(taxa, perc)), fill = "dodgerblue4", width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     expand = c(0,0), 
                     limits = c(0, 23))+
  scale_y_discrete(expand = expansion(add = c(0.5,0.5))) + 
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 12)) + 
  geom_text(aes(x = perc, y = taxa, label = taxa),
    hjust = 0,
    nudge_x = 0.3,
    colour = "dodgerblue4",
    family = "Econ Sans Cnd",
    size = 3.5)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 13)
  )

taxa_bar
```

#### Study species

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 12

species_Gs <- codes_LU %>%
  gather(col, sp_name, species1:species16) %>%
  subset(select = sp_name)

spp_Gs_s <- species_Gs %>%
  group_by(sp_name) %>%
  filter(sp_name != "NA") %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(perc = count/sum(count)*100)

sp_bar <- ggplot(spp_Gs_s) +
  geom_bar(aes(x = perc, y = reorder(sp_name, perc)), fill = "dodgerblue4", width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     expand = c(0,0), 
                     limits = c(0, 23))+
  scale_y_discrete(expand = expansion(add = c(0.5,0.5))) + 
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 12)) + 
  geom_text(aes(x = perc, y = sp_name, label = sp_name),
    hjust = 0,
    nudge_x = 0.3,
    colour = "dodgerblue4",
    family = "Econ Sans Cnd",
    size = 3.5)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 13)
  )

sp_bar
```

#### Conservation problem being addressed

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 7

consv <- codes_LU %>%
  gather(col, consv_topic, reason1:reason3) %>%
  subset(select = consv_topic)

consv_s <- consv %>%
  group_by(consv_topic) %>%
  filter(consv_topic != "NA") %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(perc = count/sum(count)*100)

consv_bar <- ggplot(consv_s) +
  geom_bar(aes(x = perc, y = reorder(consv_topic, perc)), fill = "dodgerblue4", width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     expand = c(0,0), 
                     limits = c(0, 18))+
  scale_y_discrete(expand = expansion(add = c(0,0))) + 
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 12)) + 
  geom_text(aes(x = perc, y = consv_topic, label = consv_topic),
    hjust = 0,
    nudge_x = 0.3,
    colour = "dodgerblue4",
    family = "Econ Sans Cnd",
    size = 3.5)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 13)
  )

consv_bar
```
:::

### Code appearance frequencies across articles (n = 118)

The proportions reported in these figures represent the percent of papers that mentioned a subcode within these broader code categories at least once.

::: panel-tabset
#### Broader categories

codes_vars2 = codes_vars
colnames(codes_vars2) <- gsub(x = colnames(codes_vars2), pattern = "\\.", replacement = " ")

  #calculate frequency of bridges vs. barriers being mentioned in articles -- need to add biotic and abiotic indicators cols
code_freq_new <- codes_vars2 %>%
  mutate(br_pos = as.numeric(if_any(c(33:48, 59:60), ~. == 1)), 
         ba_pos = as.numeric(if_any(49:58, ~. == 1)),
         br_dat = as.numeric(if_any(c("Management scale aligns with system needs", "Target species data availability", "Access to advanced technology and techniques", "Data transparency", "Environmental data availability", "Resource use data availability"), ~. == 1)), 
         br_eco = as.numeric(if_any(c("Expected climate change induced range shifts", "Distinct habitat preferences", "Highly mobile target species", "Life history information available for target species", "Fine scale management appropriate for target species", "Biotic indicators of target species presence", "Abiotic indicators of target species presence"), ~. == 1)), 
         br_soc = as.numeric(if_any(c("Resource availability", "Pre existing conditions", "Adoption incentives", "Feasibility", "Co management and communication"), ~. == 1)), 
         ba_dat = as.numeric(if_any(c("Gaps and inaccuracy", "Management scales do not align with system needs", "Model abuse", "Climate change adding uncertainty to target species habitat use"), ~. == 1)), 
         ba_eco = as.numeric(if_any(c("Need for multispecies management"), ~. == 1)), 
         ba_soc = as.numeric(if_any(c("Cost", "Indirect socioeconomic consequences", "Discontinuity", "Poor communication and outreach", "Poor program implementation"), ~. == 1)))

brba_sum <- code_freq_new %>%
  summarise(n_br = sum(br_pos), freq_br = n_br/118,
            n_ba = sum(ba_pos), freq_ba = n_ba/118,
            n_br_e = sum(br_eco), freq_br_e = n_br_e/118,
            n_br_d = sum(br_dat), freq_br_d = n_br_d/118,
            n_br_s = sum(br_soc), freq_br_s = n_br_s/118,
            n_ba_e = sum(ba_eco), freq_ba_e = n_ba_e/118,
            n_ba_d = sum(ba_dat), freq_ba_d = n_ba_d/118,
            n_ba_s = sum(ba_soc), freq_ba_s = n_ba_s/118) %>%
  gather(br_ba, value, c(2,4,6, 8, 10, 12, 14, 16)) %>%
  subset(select = c(br_ba, value)) %>%
  mutate(percent_label = scales::percent(value, accuracy = 1), 
         percent = if_else(br_ba == "freq_ba" | br_ba == "freq_ba_e" | br_ba == "freq_ba_d" | br_ba == "freq_ba_s", value, -value),
         code_level = if_else(br_ba %in% c("freq_ba", "freq_br"), "Overall bridge or barrier", 
                              if_else(br_ba %in% c("freq_ba_e", "freq_br_e"), "Ecological bridge or barrier", 
                              if_else(br_ba %in% c("freq_ba_d", "freq_br_d"), "Data bridge or barrier",
                              if_else(br_ba %in% c("freq_ba_s", "freq_br_s"), "Social bridge or barrier", NA)))),
         code_brba = if_else(br_ba %in% c("freq_br", "freq_br_e", "freq_br_s", "freq_br_d"), "bridge", "barrier"),
         alpha_brba = if_else(br_ba %in% c("freq_br_e", "freq_br_s", "freq_br_d", "freq_ba_e", "freq_ba_s", "freq_ba_d"), "alpha", "no_alpha")) %>%
  group_by(code_level)

brba_sum$code_level <- factor(brba_sum$code_level, levels = c("Social bridge or barrier", "Ecological bridge or barrier", "Data bridge or barrier", "Overall bridge or barrier"))
brba_sum$code_brba <- factor(brba_sum$code_brba, levels = c("bridge", "barrier"))

ggplot(brba_sum, aes(x = code_level, 
             y = percent,
             fill = code_brba)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = percent_label),
            position = position_stack(vjust = 0.65),
            color = "black", 
            size = 5) +
  ylim(-1,1)+
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"), labels = c("Bridge frequency", "Barrier frequency")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x =element_blank(),
        axis.text.y = element_text(size = 20, color = "black"),
        panel.grid = element_blank(),
        legend.position = c(-0.21, 0.99),
        legend.text = element_text(size = 16, color = "black"),
        strip.text.y.right = element_text(angle = 0, size = 16)) +
  geom_hline(yintercept = 0.0, linewidth = 1) + 
  labs(fill = "")+ 
  guides(fill = guide_legend(nrow = 1))
```

#### Data frequencies

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

  #calculate frequency of bridges vs. barriers being mentioned in articles -- need to add biotic and abiotic indicators cols
data_freq <- codes_vars2 %>%
  subset(select = c("Management scale aligns with system needs", "Target species data availability", "Access to advanced technology and techniques", "Data transparency", "Environmental data availability", "Resource use data availability", "Gaps and inaccuracy", "Management scales do not align with system needs", "Model abuse", "Climate change adding uncertainty to target species habitat use")) %>%
  gather(category, pres_abs, 1:10) %>%
  mutate(br_ba = ifelse(category == "Gaps and inaccuracy" | category == "Management scales do not align with system needs" | category == "Model abuse" | category == "Climate change adding uncertainty to target species habitat use", "Data barrier", "Data bridge")) %>%
  group_by(category, br_ba)  %>%
  summarise(n = sum(pres_abs), 
            freq = (n/118)) %>%
  mutate(percent_label = scales::percent(freq, accuracy = 2)) %>%
  group_by(category)

my_labs <- c("0", "25", "50", "75")
data_freq$br_ba <- factor(as.factor(data_freq$br_ba), levels = c("Data bridge", "Data barrier"))

data_bar <- ggplot(data_freq) +
  geom_bar(aes(x = freq, y = reorder(category, freq), fill = br_ba), width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     limits = c(0,0.9),
                     expand = c(0,0), 
                     breaks = c(0, 0.25, 0.5, 0.75),
                     labels = my_labs)+
  scale_y_discrete(expand = expansion(add = c(0,0))) +
  tidyquant::theme_tq() +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 16), 
    axis.text.y = element_text(family = "Econ Sans Cnd", size = 16), 
    legend.position = "none", 
    strip.text = element_text(size = 16)) + 
  geom_text(aes(x = freq, y = category, label = percent_label),
    hjust = 0,
    nudge_x = -0.0001,
    colour = "black",
    family = "Econ Sans Cnd",
    size = 6)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 20)
  ) + 
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"))+
  facet_wrap(~br_ba, scales = "free", nrow = 2)
data_bar

#### Ecological frequencies

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

  #calculate frequency of bridges vs. barriers being mentioned in articles -- need to add biotic and abiotic indicators cols
ecol_freq <- codes_vars2 %>%
  subset(select = c("Expected climate change induced range shifts", "Distinct habitat preferences", "Highly mobile target species", "Life history information available for target species", "Fine scale management appropriate for target species", "Biotic indicators of target species presence", "Abiotic indicators of target species presence","Need for multispecies management")) %>%
  gather(category, pres_abs, 1:8) %>%
  mutate(br_ba = ifelse(category == "Need for multispecies management", "Ecological barrier", "Ecological bridge")) %>%
  group_by(category, br_ba)  %>%
  summarise(n = sum(pres_abs), 
            freq = (n/118)) %>%
  mutate(percent_label = scales::percent(freq, accuracy = 2)) %>%
  group_by(category)

ecol_freq$category <- replace(ecol_freq$category, ecol_freq$category == "Fine scale management appropriate for target species", "Fine-scale management appropriate for target species")
ecol_freq$br_ba <- factor(as.factor(ecol_freq$br_ba), levels = c("Ecological bridge", "Ecological barrier"))

ecol_bar <- ggplot(ecol_freq) +
  geom_bar(aes(x = freq, y = reorder(category, freq), fill = br_ba), width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     limits = c(0,0.9),
                     expand = c(0,0), 
                     breaks = c(0, 0.25, 0.5, 0.75),
                     labels = my_labs)+
  scale_y_discrete(expand = expansion(add = c(0,0))) +
  tidyquant::theme_tq() +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 16), 
    axis.text.y = element_text(family = "Econ Sans Cnd", size = 16), 
    legend.position = "none", 
    strip.text = element_text(size = 16)) + 
  geom_text(aes(x = freq, y = category, label = percent_label),
    hjust = 0,
    nudge_x = -0.0001,
    colour = "black",
    family = "Econ Sans Cnd",
    size = 6)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 20)
  ) + 
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"))+
  facet_wrap(~br_ba, scales = "free", nrow = 2)

ecol_bar

#### Social frequencies

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

  #calculate frequency of bridges vs. barriers being mentioned in articles -- need to add biotic and abiotic indicators cols
soc_freq <- codes_vars2 %>%
  subset(select = c("Resource availability", "Pre existing conditions", "Adoption incentives", "Feasibility", "Co management and communication", "Indirect socioeconomic consequences", "Discontinuity", "Cost", "Poor communication and outreach", "Poor program implementation")) %>%
  gather(category, pres_abs, 1:9) %>%
  mutate(br_ba = ifelse(category == "Indirect socioeconomic consequences" | category == "Discontinuity" | category == "Poor communication and outreach" | category == "Poor program implementation" | category == "Cost", "Social barrier", "Social bridge")) %>%
  group_by(category, br_ba)  %>%
  summarise(n = sum(pres_abs), 
            freq = (n/118)) %>%
  mutate(percent_label = scales::percent(freq, accuracy = 2)) %>%
  group_by(category)

soc_freq$category <- replace(soc_freq$category, soc_freq$category == "Co management and communication", "Co-management and communication")
soc_freq$category <- replace(soc_freq$category, soc_freq$category == "Pre existing conditions", "Pre-existing conditions")
soc_freq$br_ba <- factor(as.factor(soc_freq$br_ba), levels = c("Social bridge", "Social barrier"))

soc_bar <- ggplot(soc_freq) +
  geom_bar(aes(x = freq, y = reorder(category, freq), fill = br_ba), width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     limits = c(0,0.9),
                     expand = c(0,0), 
                     breaks = c(0, 0.25, 0.5, 0.75),
                     labels = my_labs)+
  scale_y_discrete(expand = expansion(add = c(0,0))) +
  tidyquant::theme_tq() +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 16), 
    axis.text.y = element_text(family = "Econ Sans Cnd", size = 16), 
    legend.position = "none", 
    strip.text = element_text(size = 16)) + 
  geom_text(aes(x = freq, y = category, label = percent_label),
    hjust = 0,
    nudge_x = -0.0001,
    colour = "black",
    family = "Econ Sans Cnd",
    size = 6)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 20)
  ) + 
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"))+
  facet_wrap(~br_ba, scales = "free", nrow = 2)
soc_bar

@@

### Code appearance frequencies within code categories

Here, I have plotted the same information as the previous section, but as a percentage of the overall bridges and barriers groupings. For example, data bridges make up 37% of all bridge codes. The same pattern holds for the other sub-panels. For example, on the ecological page, abiotic indicators account for 22% of all ecological bridge codes.

::: panel-tabset
#### Broader categories

@@

@@

#### Data codes

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

data_freq <- code_freq %>%
  filter(code_cat == "Data bridge" | code_cat == "Data barrier")

>>>>

---
title: "DOM bridges and barriers: systematic review and survey"
author: "Emily Nazario | UC Santa Cruz | enazario'@'ucsc.edu"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
editor_options: 
  chunk_output_type: console
---

@@

## Questions

**Which themes overlap between the literature and surveys?**

**Which themes only appear in the literature? Which themes are unique to the surveys?**

**How do themes vary between fishing vs. shipping contexts?**

**How do themes vary with positionality of the respondent (surveys only)?**

## Result 1: Bridges and barriers according to review and survey

Here I've presented three options for presenting the themes and sub-themes identified in the review and surveys. I'd love feedback on which option or combination of options best visualize the codes! Depending on your feedback, I'll then spend time cleaning up the best option.

### Option 1 - Sankey diagram

Plot below only presents the review themes and sub-themes. If this is the best way to communicate the codes, I will add a diagram visualizing the survey codes.

@@

  xlim(0.89, 3.6) +
  ylim(-1, 100)+
  labs(y = "Proportion of all codes (%)", x = "") +
  theme_minimal()+
  theme(axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16), 
        legend.position = "none", 
        axis.text.x = element_blank(), 
        panel.grid.major = element_blank())+
  scale_fill_manual(values = c("#F48136", "#F46036", "#F44336", "#CFE2F3", "#C5E0B4", "#45818E"))+
  scale_color_manual(values = c("#F48136", "#F46036", "#F44336", "#CFE2F3", "#C5E0B4", "#45818E"))
```

### Option 2 - Sunburst figure

Plots below are from an old draft of the review themes and sub-themes. If this is the nicest way to plot the codes, I will update the diagram and generate one for the survey themes as well.

![](images/barriers_old.png){fig-align="center" width="70%"}

### Option 3 - Table

::: panel-tabset

#### Review: Bridge themes

#### Review: Barrier themes

#### Survey: Bridge themes

bri_codes_s <- code_freq_s %>% 
  filter(bridge_barr == "Bridge") %>% 
  subset(select = -c(total, bridge_barr))

colnames(bri_codes_s) <- c("Code category", "Code label")

bri_codes_s %>%
  kbl() %>%
  kable_styling()

@@

#### Survey: Barrier themes

bar_codes_s <- code_freq_s %>% 
  filter(bridge_barr == "Barrier") %>% 
  subset(select = -c(total, bridge_barr))

colnames(bar_codes_s) <- c("Code category", "Code label")

bar_codes_s %>%
  kbl() %>%
  kable_styling()

@@

## Result 2: Frequencies according to review and survey

::: panel-tabset

### Literature 
Code appearance frequencies across articles (n = 118). The proportions reported in these figures represent the percent of papers that mentioned a subcode within these broader code categories at least once.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center
#| include: FALSE

@@

#ggsave("figures/ms_figs/all_freq/all_freq.png", width = 8, height = 5, units = ("in"))
```

#reformat data
code_sum_br <- code_freq %>%
  filter(bridge_barr == "Bridge") %>%
  mutate(n_total = sum(total)) %>%
  group_by(code_cat, code_name) %>%
  summarise(value = total/n_total) %>% #percentages represent total number of bridge or barriers codes that code represents
  mutate(percent = scales::percent(value, accuracy = 1), 
         code_name = factor(code_name)) %>%
  ungroup() %>%
  group_by(code_cat) %>%
  arrange(code_name, desc(percent))

#plot
br_freq <- ggplot(code_sum_br, aes(x = value*100, y = reorder(code_name, value))) + 
  geom_point(size = 5, color = "goldenrod") + 
  scale_x_continuous(limits = c(0, 30), 
                     breaks = seq(0, 30, 5), 
                     name = "Percent appearance (%)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 0), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

#ggsave("figures/ms_figs/all_freq/br_freq.png", width = 7, height = 8, units = c("in"))

@@

#reformat data
code_sum_ba <- code_freq %>%
  filter(bridge_barr == "Barrier") %>%
  mutate(n_total = sum(total)) %>%
  group_by(code_cat, code_name) %>%
  summarise(value = total/n_total) %>% #percentages represent total number of bridge or barriers codes that code represents
  mutate(percent = scales::percent(value, accuracy = 1), 
         code_name = factor(code_name)) %>%
  ungroup() %>%
  group_by(code_cat) %>%
  arrange(code_name, desc(percent))

#plot
ba_freq <- ggplot(code_sum_ba, aes(x = value*100, y = reorder(code_name, value))) + 
  geom_point(size = 5, color = "lightskyblue3") + 
  scale_x_continuous(limits = c(0, 30), 
                     breaks = seq(0, 30, 5), 
                     name = "Percent appearance (%)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 0), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

#ggsave("figures/ms_figs/all_freq/ba_freq.png", width = 8, height = 8, units = c("in"))
